---
title: "In-Class: Flu Shot Coverage"
format:
  html:
    toc: true
    code-tools: true
    embed-resources: true
execute:
  echo: true
  warning: true
---

## Setup

```{r}
#install.packages(c("RMariaDB"))
library(RMariaDB)

con <- dbConnect(
  RMariaDB::MariaDB(),
  host     = "hank.test",
  dbname   = "synthea",
  user     = "root",
  password = "",
  port     = 3306
)
```

::: {.callout-note}
Today's queries all use `patients` and `immunizations`. Before writing
anything, take a minute to look at what's actually in these tables. Clicking
the magnifying glass, using `DESCRIBE immunizations;` and a quick `SELECT * FROM immunizations LIMIT 10;` will save you guessing at column names.
:::

## Warm-up: what immunizations are on record?

```{r}
#| label: warmup-immunization-counts
#| connection: con

# TODO: count how many records exist per immunization DESCRIPTION,
# most common first. Same GROUP BY + COUNT(*) pattern you already know.
```

## Step 1: patients who've had a flu shot

Same shape as "patients who have a hypertension diagnosis" from the
lecture. A yes/no question about each patient, answered with `EXISTS`.

```{r}
#| label: patients-with-flu-shot
#| connection: con

# TODO: use EXISTS to find patients with at least one immunization
# record whose DESCRIPTION contains "Influenza".
```

## Step 2: who has, and who hasn't

Now the "with and without" version in one query. This is the same shape 
as the hypertension file's "patients with and without hypertension" section.

```{r}
#| label: flu-shot-status-naive
#| connection: con

# TODO: LEFT JOIN patients to immunizations (matched on the same
# Influenza filter as Step 1), and use CASE WHEN to label each patient
# "Immunized" or "Not immunized". Don't add DISTINCT yet -- look closely
# at the result first.
```

**Before you move on:** does every patient appear in your result exactly
once? If not, why not? (Hint: think back to why the hypertension file's
first "patients with hypertension" query could return the same patient
twice.)


## Step 3: define "immunized" as up to date, not just ever vaccinated

A flu shot from 2015 shouldn't count as current. Revise your query so
that a patient only counts as `"Immunized"` if their **most recent**
flu shot was after 2020 — and include that most-recent date as its own
column, so the result shows its own evidence rather than just a label.

```{r}
#| label: flu-shot-status-recent
#| connection: con

# TODO: rewrite Step 2's query using GROUP BY instead of a plain LEFT JOIN, so you can pull MAX(i.DATE) per patient as last_flu_shot_date. Then base the CASE WHEN on whether that # date is after 2020, not just on whether a matching row exists.

# Think about what MAX(i.DATE) actually returns for a patient # with NO flu shot at all -- and why comparing that result to a # date still gives you the right answer in your CASE WHEN, # # without needing a separate check for "no record."
```

## Cleanup

```{r}
#| label: disconnect

dbDisconnect(con)
```
