---
title: "Solution: Flu Shot Coverage"
format:
  html:
    toc: true
    code-tools: true
    embed-resources: true
execute:
  echo: true
  warning: true
---

## Setup

```{r}
library(RMariaDB)

con <- dbConnect(
  RMariaDB::MariaDB(),
  host     = "hank.test",
  dbname   = "synthea",
  user     = "root",
  password = "",
  port     = 3306
)
```

## Warm-up: what immunizations are on record?

```{r}
#| label: warmup-immunization-counts
#| connection: con

immunization_counts <- dbGetQuery(con, "
SELECT DESCRIPTION, COUNT(*) AS n
FROM immunizations
GROUP BY DESCRIPTION
ORDER BY n DESC")
View(immunization_counts)

```

## Step 1: patients who've had a flu shot

`EXISTS` answers a yes/no question per patient — it never actually
returns rows from `immunizations`, so a patient with five flu shots on
record still appears exactly once here.

```{r}
#| label: patients-with-flu-shot
#| connection: con

patients_with_flu_shot <- dbGetQuery(con, "
SELECT p.Id, p.LAST
FROM patients p
WHERE EXISTS (
    SELECT 1 FROM immunizations i
    WHERE i.PATIENT = p.Id AND i.DESCRIPTION LIKE '%Influenza%'
)")
View(patients_with_flu_shot)
```

## Step 2: who has, and who hasn't

### First pass — watch what happens

```{r}
#| label: flu-shot-status-naive
#| connection: con

flu_shot_status_naive <- dbGetQuery(con, "
SELECT
    p.Id, p.LAST,
    CASE WHEN i.PATIENT IS NOT NULL THEN 'Immunized' ELSE 'Not immunized' END AS flu_status
FROM
    patients p
    LEFT JOIN immunizations i ON p.Id = i.PATIENT AND i.DESCRIPTION LIKE '%Influenza%'
ORDER BY p.Id")
View(flu_shot_status_naive)
```

A patient with **more than one** flu shot on record (a real-world
pattern — flu vaccines are typically given annually) shows up more than
once here, once per matching immunization row. This is exactly the same
duplication issue as the hypertension file's very first "patients with
hypertension" query: a `JOIN` (or `LEFT JOIN`) returns one row per
*match*, not one row per patient, whenever the matched table can have
more than one row per patient.

### Last Flu Shot
```{r}
#| label: flu-shot-status-recent
#| connection: con

flu_shot_status <- dbGetQuery(con, "
SELECT
    p.Id, p.LAST,
    CASE WHEN MAX(i.DATE) > '2020-01-01' THEN 'Immunized' ELSE 'Not immunized' END AS flu_status,
    MAX(i.DATE) AS last_flu_shot_date
FROM
    patients p
    LEFT JOIN immunizations i ON p.Id = i.PATIENT AND i.DESCRIPTION LIKE '%Influenza%'
GROUP BY p.Id, p.LAST
ORDER BY flu_status, p.Id")
View(flu_shot_status)
```

```{r}
#| label: another solution
#| connection: con
flu_shot_status_b <- dbGetQuery(con, "
Select
	b.id
	, b.LAST
	, (case	
		when b.DESCRIPTION is null then 'Not Recently Immunized'
		else 'Recently Immunized'
		END) AS flu_status
	, b.DATE
FROM
	(select
		p.id
		, p.last
		, p.FIRST
		, i.DESCRIPTION
		, i.DATE
		, ROW_NUMBER() over (Partition By p.id order by i.Date DESC) rownum
	from
		patients p
		left join immunizations i
			on p.Id = i.PATIENT
			and lower(i.description) like '%influ%'
			and i.DATE > '2020-01-01') b
WHERE
	b.rownum = 1
ORDER BY
	b.DATE
	, flu_status
	, b.Id
")
```

## Cleanup

```{r}
#| label: disconnect

dbDisconnect(con)

```