---
title: "In-Class Exercise: HbA1c Control Trends — Clinics A & C"
format:
  html:
    toc: true
    code-tools: true
execute:
  echo: true
  warning: true
---

## Background

You're a junior data analyst at a healthcare system. Department
leadership has asked your team for an analysis of **trends in HbA1c
control at four clinics** — is control improving, holding steady, or
slipping, clinic by clinic?

Your data engineer pulled the data, and — kind of lazily — handed you
**three separate files** instead of one clean extract:

- `encounters.csv` — one row per patient encounter, but instead of the
  clinic name it only lists `provider_id` (apparently linking encounters
  to clinics was somebody else's job)
- `labs.csv` — one row per lab result
- `provider_lookup.csv` — the file that actually maps each `provider_id`
  to a clinic, which someone should have joined in before handing this
  to you

Before you got to the data, your senior colleague started this
document to lay out what's needed — but only got as far as comments.
The code is yours to write. Today you'll get **Clinics A and C** done;
the other two are this week's homework.

## What "in control" means here

> **Denominator (per clinic, per month):** number of distinct patients
> with at least one encounter at that clinic that month.
>
> **Numerator (per clinic, per month):** of those, how many have an
> `HbA1c` lab result recorded **that same month** below 8%. A patient
> with no HbA1c result that month does not count — a missing test isn't
> the same as good control.

These files are not clean. Budget real time for step 2.

```{r}
#| label: setup
#| message: false

library(readr)
library(dplyr)
library(lubridate)
library(qicharts2)
```

## 1. Load and look

```{r}
#| label: load-data
#| message: false

encounters <- read_csv("data/encounters.csv")
labs <- read_csv("data/labs.csv")
lookup <- read_csv("data/provider_lookup.csv")
```

```{r}
#| label: explore

glimpse(encounters)
glimpse(labs)
glimpse(lookup)
```

Notice `encounters` has `provider_id`, not `clinic_id` — that's the
"three files instead of one" problem in a nutshell. Look closely at the
date columns in both files, and at `result_value` in `labs`, before you
go further.

## 2. Clean the dates

```{r}
#| label: parse-dates

# TODO: write a helper that parses a character vector containing a MIX
# of "YYYY-MM-DD" and "MM/DD/YYYY" strings into a single Date column.
# Apply it to encounter_date and result_date.
#
# Sanity check afterward: zero NAs in either parsed date column.
```

## 3. Attach clinic_id to every encounter

```{r}
#| label: crosswalk-join

# TODO: join `encounters` to `provider_lookup` on provider_id, so every
# encounter row gets the right clinic_id. Confirm zero unmatched rows
# before moving on -- if you see NAs in clinic_id after the join, don't
# just drop them; find out why they didn't match.
```

## 4. Build the monthly denominator

```{r}
#| label: denominator

# TODO: for Clinics A and C, count DISTINCT patients per clinic per
# month. Watch out -- a patient can have more than one encounter in the
# same month, so a plain row count will overcount.
#
# Hint: lubridate::floor_date(date, "month") turns any date into the
# first of its month, a convenient grouping key.
```

## 5. Build the monthly numerator

```{r}
#| label: numerator

# TODO: filter `labs` to test_name == "HbA1c". Convert result_value to
# numeric -- what happens to entries that aren't valid numbers? Remove
# exact duplicate rows. Then, for each clinic and month, count DISTINCT
# patients with a qualifying (< 8) result recorded that month.
#
# labs alone don't carry clinic_id -- you'll need the joined encounter
# table from step 3 to know which clinic-month a given result belongs
# to.
```

## 6. Put it together

```{r}
#| label: assemble

# TODO: combine your denominator and numerator tables into one data
# frame with columns clinic_id, month, denominator, numerator, filtered
# to Clinic A and Clinic C. A clinic-month with no qualifying result
# should show numerator = 0, not a missing row.
```

## 7. Build the run charts

```{r}
#| label: run-chart-a

# TODO: filter to Clinic A and pass to qic() with chart = "run". A run
# chart plots the median as a flat reference line and applies the
# Anhøj runs rules (unusually long runs / unusually few crossings) --
# it does not use 3-sigma control limits the way a p-chart does.
```

```{r}
#| label: run-chart-c

# TODO: repeat for Clinic C.
```

## 8. Interpret, and draft a one-line status for leadership

For each clinic, inspect the chart object's underlying data
(`your_object$data`) for the `runs.signal` column, and write 2-3
sentences plus a one-line summary you could drop into a status update.

*Clinic A:*

> [Is there a signal? What does the pattern of points look like?]
> **One-line status:** [e.g., "Clinic A: stable, no action needed."]

*Clinic C:*

> [Is there a signal? Does a run chart's "unusually long run / unusually
> few crossings" logic seem well-suited to whatever pattern you see
> here, or can you imagine a pattern it would miss?]

> **One-line status:** [your line here]
