---
title: "Solution: HbA1c Control Trends — Clinics A & C"
format:
  html:
    toc: true
    code-tools: true
    embed-resources: true
execute:
  echo: true
  warning: true
---

## Setup

```{r}
#| label: setup
#| message: false

library(readr)
library(dplyr)
library(lubridate)
library(qicharts2)
```

## 1. Load and look

```{r}
#| label: load-data
#| message: false

encounters <- read_csv("data/encounters.csv")
labs <- read_csv("data/labs.csv")
lookup <- read_csv("data/provider_lookup.csv")
glimpse(encounters)
glimpse(lookup)
```

`encounter_date` and `result_date` are both `<chr>` — a mix of
`YYYY-MM-DD` and `MM/DD/YYYY` strings. `encounters` has `provider_id`,
not `clinic_id`; `provider_lookup` is the crosswalk that supplies it.

## 2. Clean the dates

```{r}
#| label: parse-dates

parse_mixed_date <- function(x) {
  d_iso <- suppressWarnings(as.Date(x, format = "%Y-%m-%d"))
  d_us  <- suppressWarnings(as.Date(x, format = "%m/%d/%Y"))
  coalesce(d_iso, d_us)
}

encounters <- encounters %>% mutate(encounter_date = parse_mixed_date(encounter_date))
labs <- labs %>% mutate(result_date = parse_mixed_date(result_date))

sum(is.na(encounters$encounter_date))
sum(is.na(labs$result_date))
```

## 3. Attach clinic_id to every encounter

```{r}
#| label: crosswalk-join

encounters <- encounters %>% left_join(lookup, by = "provider_id")

sum(is.na(encounters$clinic_id))
```

274 rows come out unmatched — but none of them belong to Clinic A or
Clinic C. That problem is confined entirely to Clinics B and D, whose
source data has a provider-ID formatting issue A and C don't have (more
on that in the homework). Confirm that for yourself before moving on:

```{r}
#| label: crosswalk-check-ac

table(encounters$clinic_id, useNA = "ifany")
```

## 4. Build the monthly denominator

```{r}
#| label: denominator

denom <- encounters %>%
  filter(clinic_id %in% c("Clinic A", "Clinic C")) %>%
  mutate(month = floor_date(encounter_date, "month")) %>%
  distinct(clinic_id, month, patient_id) %>%
  count(clinic_id, month, name = "denominator")

head(denom)
```

`distinct()` before `count()` is the key step — a patient can have 1-3
encounters in the same month, and counting raw rows would overcount.

## 5. Build the monthly numerator

```{r}
#| label: numerator

hba1c_clean <- labs %>%
  filter(test_name == "HbA1c") %>%
  distinct(patient_id, result_date, test_name, result_value) %>%
  mutate(result_value = suppressWarnings(as.numeric(result_value))) %>%
  filter(!is.na(result_value)) %>%
  mutate(month = floor_date(result_date, "month"))

patient_clinic_month <- encounters %>%
  filter(clinic_id %in% c("Clinic A", "Clinic C")) %>%
  mutate(month = floor_date(encounter_date, "month")) %>%
  distinct(clinic_id, month, patient_id)

numer <- hba1c_clean %>%
  filter(result_value < 8) %>%
  inner_join(patient_clinic_month, by = c("patient_id", "month")) %>%
  distinct(clinic_id, month, patient_id) %>%
  count(clinic_id, month, name = "numerator")

head(numer)
```

## 6. Put it together

```{r}
#| label: assemble

measure <- denom %>%
  left_join(numer, by = c("clinic_id", "month")) %>%
  mutate(numerator = coalesce(numerator, 0L))

head(measure)
```

## 7. Build the run charts

```{r}
#| label: run-chart-a

q_a <- qic(
  x = month, y = numerator, n = denominator,
  data = measure %>% filter(clinic_id == "Clinic A") %>% arrange(month),
  chart = "run", title = "Clinic A"
)
q_a
```

```{r}
#| label: run-chart-c

q_c <- qic(
  x = month, y = numerator, n = denominator,
  data = measure %>% filter(clinic_id == "Clinic C") %>% arrange(month),
  chart = "run", title = "Clinic C"
)
q_c
```

## 8. Interpret, and a status line for leadership

```{r}
#| label: signals

any(q_a$data$runs.signal, na.rm = TRUE)
any(q_c$data$runs.signal, na.rm = TRUE)
```

*Clinic A:*

No runs signal. The monthly rate bounces around a flat median with no
unusually long run on one side and no unusually thin crossing count —
the picture of ordinary common-cause variation.

 **One-line status:** Clinic A — stable, no signal, no action needed.

*Clinic C:*

Also **no runs signal** on the run chart. But Clinic C is not actually 
a "nothing happened" clinic. One specific month (September 2024) has a 
rate far below every other month, an isolated dip that snaps right back 
the next month. A run chart tests for *sustained* non-random patterns.
A single extreme point, surrounded by ordinary values on both sides, 
doesn't create a long run or thin out the crossing count enough to trip 
either rule. **This is a real limitation of run charts**: well-suited to 
persistent shifts, but capable of missing a single dramatic outlier entirely.
That's exactly the gap a p-chart's 3-sigma control limits are built to
catch. Keep this in mind: the homework uses `qic(chart = "p")`
instead of plain run charts, specifically so isolated points don't
slip through.

**One-line status:** Clinic C — no signal detected by this chart type,
but a closer look (control chart) is warranted before reporting "all
clear" to leadership.
