---
title: "Homework: HbA1c Control Trends — Clinics B & D, and a Simple ITS v2"
format:
  html:
    toc: true
    code-tools: true
    embed-resources: true
execute:
  echo: true
  warning: true
---

## Setup

```{r}
#| label: setup
#| message: false

library(readr)
library(dplyr)
library(lubridate)
library(qicharts2)
```

## Load the data

```{r}
#| label: load-data
#| message: false

encounters <- read_csv("data/encounters.csv")
labs <- read_csv("data/labs.csv")
lookup <- read_csv("data/provider_lookup.csv")
```

Same three files as class. This time you're finishing Clinics B and D,
which have two additional problems A and C didn't.

## Part 1: Build the measure

### Parse dates (same as class)

```{r}
#| label: parse-dates

# TODO: reuse your mixed-date parser from class on encounter_date and
# result_date.
```

### Task 1: standardize the join key, then join

```{r}
#| label: standardize-and-join

# TODO: first, try the plain join from class:
#   encounters %>% left_join(lookup, by = "provider_id")
# and check how many rows come back with a missing clinic_id. Then look
# at a sample of the unmatched provider_id values -- what's inconsistent
# about them relative to provider_lookup.csv?
#
# TODO: create a standardized version of provider_id on BOTH tables
# (case + whitespace), then join on the standardized version. Confirm
# zero unmatched rows afterward.
```

### Build the denominator

```{r}
#| label: denominator

# TODO: same logic as class,
# filter to Clinic B and Clinic D.
# set dates to month level
# reduce to unique patients per clinic per month rows
# count the number for rows
```

### Clean the labs and handle Task 2: implausible values

```{r}
#| label: clean-labs-filter

# TODO: filter to test_name == "HbA1c", convert result_value to numeric,
# drop non-numeric/missing entries, remove exact duplicates -- same logic as
# class.

```

```{r}
#| label: clean-labs-filter


# TODO (Task 2): before computing the numerator, add a plausible-range
# filter on the numeric value (a reasonable range for this exercise is
# roughly 4-15). Report how many rows this excludes. Also convert dates to month. 

```

### Build the numerator and assemble the measure table

```{r}
#| label: numerator-patients

# TODO: same join-back-to-clinic-month logic as class
# First, identify unique patient visits in a month.
```

```{r}
#| label: numerator-labs
# Finish generating a numerator - # of patients with controlled values (under 8)
# Filter to controlled lab values, join to the patient-by month file above.
# Identify unique patients per month per clinic
# Count how many patients per month had controlled lab values. This is the numerator.
# 

```

```{r}
#| label: numerator-assembly
# Join numerator and denominator together


```


## Part 2: Control charts

```{r}
#| label: chart-b

# TODO: qic() with chart = "p" for Clinic B. Store the result.
```

*Clinic B interpretation:*

> [3-5 sentences: signal present? sigma or runs? plausible cause? next
> step?]
>
> **One-line status:** [your line here]

```{r}
#| label: chart-d

# TODO: qic() with chart = "p" for Clinic D.
```

*Clinic D interpretation:*

> [3-5 sentences.]
>
> **One-line status:** [your line here]

## Part 3: A simple interrupted time series (Clinic B)

```{r}
#| label: its-setup

# TODO: build a data frame for Clinic B with:
#   - rate = numerator / denominator
#   - t = a simple 1, 2, 3, ... month counter
#   - post = 0/1 indicator for month >= 2025-03-01
#   - t_post = post * (a counter that starts at 1 on the intervention
#     month and counts up from there, 0 before it)
```

```{r}
#| label: its-model

# TODO: fit lm(rate ~ t + post, data = ...) and look at
# summary(). Extract/interpret the post and t_post coefficients.
```

*Interpretation:*

> [What does the `post` coefficient mean, and is it significant? What
> does `t_post` mean, and is it significant? Does this agree with what
> the control chart's signal already told you? Explain, in your own
> words, what each method is actually testing and why they might (or
> might not) agree.]
